Raising the role of Vocabulary Hubs for Semantic data interoperability in Dataspaces

### 2024-06-03 EBDVF-2024/paper.qmd
start EBDVF-2024/paper.qmd

---
title: Consolidating Wind Farm Data in Data Spaces for Training Predictive Maintenance ML

abstract: |
  The UNDERPIN project addresses the challenge of training machine learning (ML) models for predictive maintenance of wind farms by leveraging consolidated training data provided by multiple partners. This collaborative effort is facilitated through a European-wide Data Space for Manufacturing for dynamic asset management and predictive and prescriptive maintenance established as part of the project. Data spaces allow trusted participants to share data while maintaining data sovereignty. However, consolidating diverse data sets requires a clear semantics and mapping framework.
  To solve this, the project employs Semantic Web standards and technologies to model and process the data. Using OWL (Web Ontology Language), vocabularies and ontologies are developed to represent the data with clear, standardized semantics. Two key software components, GraphDB and PoolParty, are utilized to implement these standards and provide a semantic layer to efficiently process the consolidated data. The main benefit is to support ML training by presenting a consolidated and uniform training data set. Our approach aims to enhance the quality and reliability of predictive maintenance models for wind farms.
  A demonstration from the UNDERPIN project showcases the process of modeling and processing the data, illustrating the practical application of these semantic technologies. This example highlights the project's approach and the potential improvements in predictive maintenance through enhanced data integration and processing.

author:
  - name: Robert David
    email: 
    orcid: 
    affiliation: 
     name: The Semantic Web Company
     country: Austria
  - name: Axel Weißenfeld
    email: 
    orcid: 
    affiliation: 
      name: Austrian Institute of Technology
      country: Austria
  - name: Petar Ivanov
    email: petar.ivanov@ontotext.com
    orcid: 0000-0001-8448-1005
  - name: Nikola Tulechki
    email: 
    orcid: 
  - name: Vladimir Alexiev
    email: vladimir.alexiev@ontotext.com
    orcid: 0000-0001-7508-7428
    affiliation: 
      name: Ontotext
      country: Bulgaria
acm-metadata:
  keywords:
    - data spaces
    - oil and gas
    - renewable energy
    - refineries
    - windfarms
    - time series
    - ontologies
    - semantic technologies
  final: true
  # acmart-options: sigconf, review
  # copyright-year: 2024
  # acm-year: 2024
  copyright: acmcopyright
  # doi: 
  # conference-acronym: 
  # conference-name: 
  # conference-date: 
  # conference-location: 
  # isbn:

>>>>

---
title: Raising the role of Vocabulary Hubs for Semantic data interoperability in Dataspaces

abstract: |
  Dataspaces are an important enabler for industrial sharing data (either commercially licensed or private).
  Europe is investing heavily into sectoral dataspaces, federation and orchestration platforms like SIMPL, Eclipse DSC, GXFS, etc.
  Still, dataspaces enable shared data access, but do not solve the data interoperability problem.
  For that, the consumer would like to see the data from different providers in a harmonized and semantically integrated data.
  The Vocabulary Hub service (part of the IDSA RAM) provides a repository for ontologies and vocabularies.
  We describe an approach of raising the role of the vocabulary hub to also allow richer metadata description (e.g. the meaning of every column in a tabular dataset),
  and binding semantic descriptions to ingested datasets, thus providing on-the-fly data semantization and easing data querying.
  This is achieved through the integration of two commercial semantic products (PoolParty and GraphDB),
  leveraging the partnership between the Semantic Web Company and Ontotext,
  and is being developed within the frame of the Digital Europe project UNDERPIN,
  with applications to refinery and wind farm data.

author:
  - name: Martin Kaltenböck
    email: martin.kaltenboeck@semantic-web.com
  - name: Robert David
    email: robert.david@semantic-web.com
    orcid: 0000-0002-3244-5341
    affiliation: 
     name: The Semantic Web Company
     country: Austria
  - name: Vladimir Alexiev
    email: vladimir.alexiev@ontotext.com
    orcid: 0000-0001-7508-7428
  - name: Petar Ivanov
    email: petar.ivanov@ontotext.com
    orcid: 0000-0001-8448-1005
    affiliation: 
      name: Ontotext
      country: Bulgaria
acm-metadata:
  keywords:
    - data spaces
    - semantic interoperability
    - semantic technologies
    - ontologies
    - vocabulary hub
    - oil and gas
    - renewable energy
    - refineries
    - windfarms
  final: true
  # acmart-options: sigconf, review
  # copyright-year: 2024
  # acm-year: 2024
  copyright: acmcopyright
  # doi: 
  # conference-acronym: 
  # conference-name: 
  # conference-date: 
  # conference-location: 
  # isbn:
